---
title: "330-HW-3"
author: "Chris Trippe"
date: "10/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. In your own words, summarize the overarching problem and any specific questions that need to be
answered using the water data. Discuss how statistical modeling will be able to answer the posed
questions.

```{r}

```

2. Using exploratory techniques (don’t actually fit a model), explore the data to assess if a simple
linear regression (SLR) model is suitable to analyze the water data. Justify your answer using any
necessary graphics and relevant summary statistics that would suggest an SLR model would be
successful at achieving the goals of the study.

```{r}

```

3. Write out (in mathematical form with greek letters) a justifiable SLR model that would help answer
the questions in problem. Provide an interpretation of each mathematical term (β parameter) included
in your model. Using the mathematical form, discuss how your model, after fitting it to the
data, will be able to answer the questions in this problem.

```{r}

```

4. Fit your model in #3 to the water data and summarize the results by displaying the fitted model in
equation form (do NOT just provide a screen shot of the R or Python output). Interpret each of the
fitted parameters in the context of the problem. Provide a plot of the data with a fitted regression
line.

```{r}

```

5. List then justify your model assumptions using appropriate graphics or summary statistics.

```{r}

```

6. Assess the fit and predictive capability of your model. Discuss on the level of your target audience
(e.g. interpret your model R2
).

```{r}

```

7. Carry out a test that there is no relationship between snowfall and runoff (i.e., write out the hypotheses,
report an appropriate p-value, and conclude in context).

```{r}

```

8. Construct 95% confidence intervals for the slope and intercept parameters and interpret these intervals
in the context of the problem.

```{r}

```

9. In a recent winter, the site only received 4.5 inches of snowfall. What do you predict will be the
associated runoff? Provide a 95% predictive interval and interpret the interval in the context of the
problem. Do you have any hesitations performing this prediction (hint: you should)? Describe these
hesitations and their potential impact on your prediction.
