---
title: "330-HW-3"
author: "Chris Trippe and Ethan Kemeny"
date: "10/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r include=FALSE}
library(tidyverse)
library(MASS)
library(lmtest)
```

1.

Droughts are a big problem in California. Stream runoff from snowfall is one source of water. A statistical model could be used to predict the expected amount of runoff in acre-feet of a river near Bishop, California by the amount of snowfall at a site in the Sierra Nevada mountains. If the amount of water entering the area is estimated, people such as engineers, planners, and policy makers will be able to do their jobs more efficiently.

```{r, echo = FALSE, fig.width = 6, fig.height = 4}
##read in data
water_data <- read.table(file = "water.txt", header = TRUE, sep = "")
```

2. Using exploratory techniques (don’t actually fit a model), explore the data to assess if a simple
linear regression (SLR) model is suitable to analyze the water data. Justify your answer using any
necessary graphics and relevant summary statistics that would suggest an SLR model would be
successful at achieving the goals of the study.

```{r, echo = FALSE, fig.width = 3, fig.height = 2}
ggplot(water_data, aes(Precip, Runoff)) + geom_point()
```


```{r, echo = FALSE}
#check correlation, is it linear?
cor(water_data$Precip, water_data$Runoff)
```

The relationship is relatively strong and positive with a correlation of 0.938. As snowfall increases the amount of runoff increases. 

```{r, echo = FALSE}
##linear model for raw data
slr_water <- lm(Runoff~Precip, water_data)
```



```{r, echo = FALSE}
#test for variance
bptest(slr_water)
```

According to the Breusch-Pagan test, the p-value is 0.3107 which means we fail to reject the null hypothesis and the residuals can be considered to have equal variance.

```{r, echo = FALSE}
#test for normality
ks.test(stdres(slr_water),"pnorm" )
```

According to the Kolmogorov-Smirnov test, the p-value is 0.621 which means we fail to reject the null hypothesis and the residuals can be considered approximately normal about the fitted line. 

```{r, echo = FALSE, fig.width = 3, fig.height = 2.5}
cook_dist <- cooks.distance(slr_water)
outliers <- which(cook_dist > 4/length(cook_dist))
plot(cook_dist, type="h", main = "Cook's Distance")
abline(h = 4/length(cook_dist), col = "red")

```

There are only two outliers, only one of which is extreme.

The only questionable assumption is independence. Temperature, and thus the amount of snowfall, is going to be similar year to year, so the amount of runoff might not be independent of the amount of runoff from the previous year. However, we feel safe enough to proceed without manipulating the given data.

3. 

$$y_i = \beta_0 + \beta_1x_i + \epsilon_i \text{ where } \epsilon_i \sim \mathcal{N}(\mu,\,\sigma^{2})  $$
$y_i$ is the amount of runoff in acre-feet of a river near Bishop, California from the $i^{th}$ year.
$x_i$ is the amount of snowfall in inches at a site in the Sierra Nevada mountains from the $i^{th}$ year.
$\beta_0$ is the y-intercept; on average, what you would expect the amount of runoff to be if the amount of snowfall in inches was 0.
$\beta_1$ is the slope coefficient; as the snowfall increases by 1 inch, the amount of runoff is expected to increase by $\beta_1$ acre-feet, on average.
$\epsilon_i$ is the residual; the error associated with the $i^{th}$ observation.

When the amount of snowfall is plugged into the the above formula, it will provide an estimated amount of runoff. This estimate will help enginners, plonners, and policy holders more effectively do their jobs.

4. 

```{r, echo = FALSE}
#Intercept and Slope values
slr_water$coefficients

```


$$\hat{y_i} = 27014.587 + 3752.486 x_i $$


```{r, echo = FALSE, fig.width = 3, fig.height = 2}
#plots new transformed slr on original scale
ggplot(water_data,aes(Precip, Runoff)) +geom_point() + geom_smooth(se = FALSE,method = "lm")
```

$\beta_0$ is the y-intercept; on average, the amount of runoff will be 27014.587 acre-feet, when the amount of snowfall in inches is 0.
$\beta_1$ is the slope coefficient; as the snowfall increases by 1 inch, the amount of runoff is expected to increase by 3752.486 acre-feet, on average.

5. List then justify your model assumptions using appropriate graphics or summary statistics.

The assumptions are that the fitted model is linear, the amount of runoff for any observation is independent of all other runoff observations, the residuals are approximately normally distributed about the fitted line, and the variance of the residuals is equal about the fitted line. 

All of these assumptions were checked in part 2. The only questionable assumption is independence which we will consider inconsequential and move forward with the analysis.

6. Assess the fit and predictive capability of your model. Discuss on the level of your target audience
(e.g. interpret your model R2
).

```{r, echo = FALSE}
n_cv <- 250 #number of cross validations
bias_vec <- rep(NA, n_cv) #vector for biases
rpmse_vec <- rep(NA, n_cv) #vector for rpmse
n_test <- round(.1 * nrow(water_data)) #size of test set

for(i in 1:n_cv){
   
  #choose which obs. to put in test set
  test_obs <- sample(1:nrow(water_data), n_test)
  
  #Split data into test and training sets
  test_set <- water_data[test_obs,]
  train_set <- water_data[-test_obs,]
  
  #Using training data to fit a model
  train_lm <- lm(Runoff~Precip, data = train_set)
  
  #Predict test set
  test_preds <- predict.lm(train_lm, newdata = test_set)
  
  #Calculate bias
  bias_vec[i] <- mean(test_preds - test_set$Runoff)
  
  #Calculate RPMSE
  rpmse_vec[i] <- sqrt(mean((test_preds - test_set$Runoff)^2))
}
```


```{r, echo=FALSE, comment=""}
paste0("R squared equals ",summary(slr_water)$r.squared) %>% writeLines()
paste0("The bias is ", mean(bias_vec)) %>% writeLines()
paste0("The root predicted mean square error is ", mean(rpmse_vec)) %>% writeLines()
```

```{r, echo = FALSE}
cat("R squared interpretation:\n")
cat(round(summary(slr_water)$r.squared*100, 1), "percent of the variation in the amount of runoff is explained away by the amount of snowfall.")
cat("Bias interpretation:")
cat("On average, the predictions for runoff are", mean(bias_vec) ,"acre-feet too high.")
cat("Root predicted mean square error interpretation:")
cat("On average, the predicted runoff amounts are off by", mean(rpmse_vec), "feet.")
```


7. Carry out a test that there is no relationship between snowfall and runoff (i.e., write out the hypotheses,
report an appropriate p-value, and conclude in context).

$H_0: \beta_1 = 0$
$H_{\alpha}: \beta_1 \ne 0$

```{r, echo = FALSE}
#gets the summary stats of the original slr, coeffucients contains the pvalues for hypothesis test in the fourth column

p_val <- summary(lm(Runoff~Precip,water_data))$coefficient[2,4]
p_val
```

Since the p-value approximates zero and is less than $\alpha =$ 0.5, we reject the null hypothesis that there is no relationship between amount of snowfall and amount of runoff.


8. Construct 95% confidence intervals for the slope and intercept parameters and interpret these intervals
in the context of the problem.

```{r, echo = FALSE}
confint(slr_water, level = 0.95)
```

We are 95% confident that a 1 inch increase in snowfall would increase runoff by between 3316.809 and 4188.162 acre-feet, on average.

We are 95% confident that when snowfall is 0 inches, runoff would be between 20513.978 and 33515.197 acre-feet.

9. In a recent winter, the site only received 4.5 inches of snowfall. What do you predict will be the
associated runoff? Provide a 95% predictive interval and interpret the interval in the context of the
problem. Do you have any hesitations performing this prediction (hint: you should)? Describe these
hesitations and their potential impact on your prediction.

```{r, echo = FALSE}
cat("Range for snowfall values:")
range(water_data$Precip)
```


```{r, echo = FALSE}
dframe <- data.frame(Precip = 4.5)
pred_values <- predict.lm(slr_water, dframe, interval = "prediction", level = 0.95)
pred_values
```

When using the fitted model, we are 95% confident that if snowfall is 4.5 inches, the amonut of runoff will be between 25254.2 and 62547.34 acre-feet. However, predicting for  4.5 inches would be extrapolating. The fitted model was developed on snowfall values that ranged from 4.6 to 33.07. We don't know how accurately our model estimates runoff amounts outside the given range. Therefore, the prediction interval cannot be guaranteed to give accurate results. 